---
title: "MVA Assignment2"
author: "Sai Madhuri"
date: "2024-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
##### Assignment2
```{r}
file_path <- "/Users/user/Downloads/KCLT.csv"
data <- read.csv(file_path)
weather_data <- read.csv("/Users/user/Downloads/KCLT.csv")
str(weather_data)
```
##### Univariate Analysis

```{r}
#How is the distribution of actual mean temperature?

library(ggplot2)
hist(weather_data$actual_mean_temp, 
     main = "average temperature distribution",
     xlab = "average temperature",
     ylab = "number of occurences",
     col = "lightblue",
     border = "red")

```

#insights from the above histogram
#From the above histogram we can understand the below points
# 1.There are no extreme values(outliers) in the data
# 2.There is a slight negative skewness(left-skewed) in the data
# 3.The temperatures are normally distributed after 60.

```{r}
#At what range most of the mean temperatures are clustered?
ggplot(weather_data, aes(x = actual_mean_temp)) +
  geom_density() + 
  labs(title = "Density Plot of mean temperature")

```

#From the density plot we got a smooth estimate of the probability density function which tells that most the average temperatures lies between the range 40-80.


##### Bivariate Analysis
```{r}
#study the relationship between mean temperature and precipitation?
ggplot(weather_data, aes(x = actual_mean_temp, y = record_precipitation)) +
  geom_point(color = "pink") +
  labs(title = "Mean temperature Vs Mean precipitation",
       x = "Average temperature",
       y = "precipitation") +  geom_smooth(method = "lm")


```

#Insights from the above visualization 
As average temperatures rise, more evaporation occurs which in turn increases the overall precipitation on earth's surface.

```{r}
#compare the ranges of minimum and maximum temperatures?
boxplot(weather_data[,3:4])
boxplot(weather_data[,7:8])


```
#Box plot is an easy way to identify mean values and the dispersion of the datasets.

```{r}
# How the precipitation is varying over time?
ggplot(weather_data, aes(x = date, y = record_precipitation))  +
  geom_point(size = 3, 
             color = "steelblue") +
  labs(y = "Precipitation", 
       x = "time",
       title = "Precipitation change over time")
```

#From the above plot it is evident that precipitation mostly stays close to it's previous day records.


```{r}
file_path <- "/Users/user/Downloads/KCLT.csv"
data <- read.csv(file_path)
weather_data <- read.csv("/Users/user/Downloads/KCLT.csv")
str(weather_data)
```
```{r}
file_path <- "/Users/user/Downloads/KCLT.csv"
data <- read.csv(file_path)
weather_data <- read.csv("/Users/user/Downloads/KCLT.csv")
str(weather_data)

```
```{r}
file_path <- "/Users/user/Downloads/KCLT.csv"
data <- read.csv(file_path)
weather_data <- read.csv("/Users/user/Downloads/KCLT.csv")
str(weather_data)

```


#Multivariate Analysis
```{r}
#Starplot
stars(weather_data)


```
#From the starplot we can see all the columns and their respective values


```{r}

#How close the recorded temperatures are related to the measured average temperatures?
pairs(weather_data[,5:8])
library(GGally)
ggscatmat(weather_data, columns=5:8)


```
#It shows there is a strong positive correlation between the recorded temperatures and the average temprature s measured.The measured values are accurate

```{r}

#Estimate the years that recorded maximum temperature?
ggplot(weather_data, aes(x=record_max_temp_year, y=record_max_temp)) + geom_point()


```

#From the above graph we can estimate the years that had record maximun temperature i.e. 1953 and 2020

```{r}

#How are the independent variables(average and recorded values) related to each other?
#Compute correlation matrix
cor_matrix <- cor(weather_data[, 5:8])
#Create a heatmap of the correlation matrix
heatmap(cor_matrix, main = "Heatmap of Correlation Matrix")


```

#To show relationship between multiple variables heatmap is useful.The pale color shows low correlation and the dark color shows high correlation among the variables.








