---
title: "MVA6_FactorAnalysis"
author: "Sai Madhuri"
date: "2024-03-22"
output: html_document
---

```{r}
#Reading data
autompg <- read.csv("/Users/user/Downloads/auto-mpg.csv")
str(autompg)
autompg$horsepower <- as.integer(autompg$horsepower)
str(autompg)
```

```{r}
# cleaning data
auto1 <- na.omit(autompg)

```

```{r}
# load library for factor analysis
library(ggplot2)
library(psych)

```


#1.Decide how many Factors are ideal for your dataset?
```{r}
data <- auto1[,1:6]
#To answer question 1 we perform Parallel analysis Scree Plots
fa.parallel(data)
```

#Output:Parallel analysis suggests that the number of factors =  2  and the number of components =  1

#From the above two factors are ideal for the autompg dataset


#2.Explain the output for your factor model?
```{r}
fit.pc <- principal(data, nfactors=2, rotate="varimax")
fit.pc
```

#High absolute values (close to 1) indicate a strong relationship between the variable and the factor.
->Example: For displacement variable, among the two factors (RC1,RC2), RC1 is having high absolute value so we can conclude that displacement variable is explained better by RC1. Like this we can analyze the relationship between the variables and the factors.

#h2 explains how much variance of the variables are explained by the factors.
->Example: For displacement variable, 0.950 of it's variance is explained by the factors 

#u2 indicates the amount of variance not explained by the factors
->Example: For displacement variable, 0.049 of it's variance is unique and not explained by the factors.

#Principal Components Analysis


Call: principal(r = data, nfactors = 2, rotate = "varimax")
Standardized loadings (pattern matrix) based upon correlation matrix

                       RC1  RC2
SS loadings           3.99 1.53
Proportion Var        0.66 0.25
Cumulative Var        0.66 0.92
Proportion Explained  0.72 0.28
Cumulative Proportion 0.72 1.00

Mean item complexity =  1.3
Test of the hypothesis that 2 components are sufficient.

The root mean square of the residuals (RMSR) is  0.03 
 with the empirical chi square  11.49  with prob <  0.022 

Fit based upon off diagonal values = 1

```{r}
round(fit.pc$values, 3)
```

```{r}
fit.pc$loadings
```
```{r}
# Communalities
fit.pc$communality

```

```{r}
# Rotated factor scores, Notice the columns ordering: RC1, RC3, RC2 and RC4
fit.pc$scores

```

```{r}
fa.plot(fit.pc) # See Correlations within Factors

```


#3.Show the columns that go into each factor?
```{r}
fa.diagram(fit.pc) # Visualize the relationship

```

#From the above we can infer that weight,displacement,cylinders,mpg are well explained by RC1 and acceleration is explained better by RC2.

#4.Perform some visualizations using the factors?

```{r}
#very simple structure visualization
vss(data)

```
#output
Very Simple Structure
Call: vss(x = data)
VSS complexity 1 achieves a maximimum of 0.97  with  1  factors
VSS complexity 2 achieves a maximimum of 0.99  with  2  factors

The Velicer MAP achieves a minimum of 0.14  with  1  factors 
BIC achieves a minimum of  97.18  with  2  factors
Sample Size adjusted BIC achieves a minimum of  109.87  with  2  factors

Statistics by number of factors


```{r}
# Computing Correlation Matrix
corrm.auto <- cor(data)
corrm.auto

```
```{r}
plot(corrm.auto)

```

```{r}
auto_pca <- prcomp(data, scale=TRUE)
summary(auto_pca)

```
```{r}
plot(auto_pca)

```
#Biplot Visualization
```{r}
biplot(fit.pc)

```

