---
title: "MVA_MR"
author: "Sai Madhuri"
date: "2024-04-12"
output: html_document
---

```{r}
file_path <- "/Users/user/Downloads/auto-mpg.csv"
data <- read.csv(file_path)
autompg <- read.csv("/Users/user/Downloads/auto-mpg.csv")
str(autompg)
```
```{r}
autompg <- read.csv("/Users/user/Downloads/auto-mpg.csv")
autompg$horsepower <- as.integer(autompg$horsepower)
cleaned_auto <- na.omit(autompg)
str(autompg)
```



### 1.Model Development
```{r}
auto<- cleaned_auto[,1:6]
# Performing multiple regression on the dataset
fit <- lm(mpg~ cylinders+displacement+horsepower+weight+acceleration, data=auto)
#show the results
summary(fit)
```

###In this step, we loaded the auto-mpg dataset and fitted a multiple regression model using the lm() function. The model predicts the "mpg" (miles per gallon) based on several predictor variables: cylinders, displacement, horsepower, weight, acceleration.

### 2.Model Acceptance
###From the summary Pr column it is evident that weight and horsepower are the two most significant variables. R-squared value explains the fact that cylinders,displacement,horsepower,weight, acceleration variables contribute to the 70% of total variance,which is not bad. p-value< 2.2e-16 is significant so the model is acceptable.

```{r}
coefficients(fit)
```

###From the above we get information about the dependent variable mpg in equation form y=b0+ b1x1 + b2x2+...+bnxn where intercept b0=4.626431e+01, and cofficients b1=-3.979284e-01,....

```{r}
confint(fit,level=0.95)
```
```{r}
fitted(fit)
```

### 3.Residual Analysis

```{r}
library(GGally)
```
```{r}
ggpairs(data=auto, title="auto-mpg")
```

```{r}
plot(fit, which=1) # Residuals vs Fitted
plot(fit, which=2) # Normal Q-Q plot

```

```{r}
residuals <- residuals(fit)
```

```{r}
#Plot residuals against fitted values to check for homoscedasticity
plot_resid_fitted <- ggplot() +
  geom_point(aes(x = fitted(fit), y = residuals)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Fitted Values", y = "Residuals",
       title = "Residuals vs Fitted Values Plot") +
  theme_minimal()
print(plot_resid_fitted)

```


#The residual vs. fitted plot is a tool used to evaluate the assumptions and adequacy of a regression model. It helps to identify whether the model adequately captures the underlying relationships in the data or if there are issues that need to be addressed.
#The plot shows a random scatter of points around zero, the model is likely appropriate.


### 4.Prediction

```{r}
predict.lm(fit, data.frame(cylinders=8, displacement=310, horsepower=170, weight=4300,
    acceleration=9) )
```


##Here when random values are given to the independent variables cylinders,displacement, horsepower, weight and acceleration, the model predicted the dependent variable mpg would be 12.79.


### 5.Model Accuracy
```{r}
#Make predictions using the model
predicted <- predict(fit, newdata = auto)
```

```{r}
#Calculating RMSE by taking the square root of the mean of the squared differences between the actual values (auto$mpg) and the predicted values (predicted)
rmse <- sqrt(mean((auto$mpg - predicted)^2))
rmse
```

#RMSE measures the average deviation of predicted from actual values.

#Conclusion: Here the moderate RMSE value 4.21 suggests that the model is having resonable accuracy but with noticeable deviations from actual values.

### Some Visualizations

```{r}
library(car)
#Nonlinearity
# component + residual plot
crPlots(fit)
```



```{r}
# plot studentized residuals vs. fitted values
library(car)
spreadLevelPlot(fit)
```


