---
title: "MVA4_PCA"
author: "Sai Madhuri"
date: "2024-03-03"
output: html_document
---

```{r}
library(ggplot2)
autompg <- read.csv("/Users/user/Downloads/auto-mpg.csv")
str(autompg)
autompg$horsepower <- as.integer(autompg$horsepower)
str(autompg)
```

```{r}
autompg <- read.csv("/Users/user/Downloads/auto-mpg.csv")
str(autompg)
```

```{r}
autompg$horsepower <- as.numeric(autompg$horsepower)
str(autompg)
```




```{r}
#dropping non-numeric data
auto_mpg <- autompg[, -7:-9]
str(auto_mpg)
```

```{r}
#Get the correlations between the variables 

cor(auto_mpg, use = "complete.obs")
```
#Question1
```{r}
#Computing Principal Components
auto_mpg <- na.omit(auto_mpg)
auto_pca <- prcomp(auto_mpg,scale=TRUE)
auto_pca

```
```{r}
summary(auto_pca)
```
#summary of auto_pca it's evident that PC1 and PC2 plays an important role as they captures the highest proportion of variance.And we can discard the rest as it contributes less to the overall variance.Vehicles with higher values on PC1 tend to have higher fuel consumption (lower mpg), more cylinders, larger engine displacement, higher horsepower, and greater weight.


```{r}
#eigen values
eigen_auto <- auto_pca$sdev^2
names(eigen_auto) <- paste("PC",1:6,sep="")
eigen_auto

```

```{r}
sumlambdas <- sum(eigen_auto)
sumlambdas

```

```{r}
propvar <- eigen_auto/sumlambdas
propvar

```

```{r}
cumvar_auto <- cumsum(propvar)
cumvar_auto

```

```{r}
matlambdas <- rbind(eigen_auto,propvar,cumvar_auto)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)

```
```{r}
summary(auto_pca)

```


```{r}
auto_pca$rotation

```
```{r}
print(auto_pca)

```

#Screeplot
```{r}
screeplot(auto_pca, type = "line")
```

#variate representation of PC1
```{r}
#Graph PC1 and PC2 using x of prcomp()
plot(auto_pca$x[,1],auto_pca$x[,2])
```

```{r}
#Let's see how much variation in the original data PC1 accounts for
pc1.var <- auto_pca$sdev^2
pc1.var
```

```{r}
#percentage of variation that PC1 accounts for
pc1.var.per <- round(pc1.var/sum(pc1.var)*100, 1)
pc1.var.per
```

```{r}
#using bar plot to visualize the above percentage
barplot(pc1.var.per, main="Scree Plot", xlab="Principal Component", ylab="Percent Variation")
```
#PC1 accounts for almost all the variation in the data this means that there is a big difference between the PC1 and PC2 clusters


#Visualization

```{r}
#Visualizing PC's and the variation

pca.data <- data.frame(Sample=rownames(auto_pca$x),
  X=auto_pca$x[,1],
  Y=auto_pca$x[,2])
pca.data
```

```{r}
#ggplot
ggplot(data=pca.data, aes(x=X, y=Y, label=".")) +
  geom_text() +
  xlab(paste("PC1 - ", pc1.var.per[1], "%", sep="")) +
  ylab(paste("PC2 - ", pc1.var.per[2], "%", sep="")) +
  theme_bw() +
  ggtitle("PCA Graph")
 
```
#PC1 accounts for 79.8% of variation in the original data
#PC2 accounts for 12.1% of variation in the original data

```{r}
#eigen values
eigen_auto <- auto_pca$sdev^2
eigen_auto
plot(log(eigen_auto), xlab = "Component number",ylab = "log(Component variance)", type="l",main = "Log(eigenvalue) diagram")

 
```

```{r}
library(corrplot)
library(factoextra)
fviz_eig(auto_pca, addlabels = TRUE)

```


```{r}
fviz_pca_var(auto_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)

```


